Tool-driven Revolutions in Archaeological Science

### 2018-09-09 analysis/paper/jcaa-paper.Rmd
added a sentence why sharing code is worth looking at
typo corrections

One notable practice that we observed during our bibliometric analysis is that among authors of articles citing R, a subset of these authors make publicly available the code and data that they used to generate the results presented in their article. We identified `r number_of_reproducible_articles` articles that include R script and data files, either in the supplementary materials or in a trustworthy data repository. Figure \@ref(fig:archaeology-reproducible-articles) shows that although this number is small, it is increasing, hinting at the emergence of a new approach to how archaeologists share their research.

These archaeology articles that include publicly available R code and data are part of a new approach to scientific communication that has also been observed in other fields. Although we are not aware of attempts to quantify this change in the same way we have for archaeology, we have observed a number of recent publications that describe and recommend code sharing in statistics [@Baumer2014], genome biology [@markowetz2015five], computational biology [@Sandve2013], biostatistics [@Peng2009], computer science [@peng2011reproducible; @Mitchell2012], applied mathematics [@LeVeque2012], speech science [@Abari2012], political science [@king1995replication; Dafoe2014], and the social sciences generally [@Miguel2014]. As part of this growing interest in using code for research we also see manifestos aimed at researchers doing any kind of quantitative work [@nosek2015promoting; @Ince2012; @Barnes2010] and articles recommending using R in undergraduate education [@Bray2014; @Eglen2009]. Much of this literature is concerned with identifying and solving problems of irreproducibility in research. This comes from a renewed interest in basic principles of the science method: that once-off results should not be trusted, instead they should be reproducible by other members of the research community. Furthermore, an article should sufficiently describe the presented research results such that a colleague can fully understand the results and how they were obtained. Many of these authors trace problems of irreproducibility to the increased complextity of computer-based analyses (but only a limited space to describe them in a journal article), and the use of mouse-driven computer programs where the researcher's analytical decisions are not recorded during the research process, and thus cannot be shared with other researchers. In our view, using R (or similar open source programming language) to conduct research might not constitute a tool-driven revolution, but publicaly sharing the code used for research, as a solution to problems of irreproducibility could enable revolutionary change, since openly shared code could speed the adoption of code throughout a research community.

>>>>

@Touchon_McCoy_2016 had identified the use of scripting languages as being a factor for easily sharing methods and code, thereby increasing the rate of development of statistical methods. One notable practice that we observed during our bibliometric analysis is that among authors of articles citing R, a subset of these authors make publicly available the code and data that they used to generate the results presented in their article. We identified `r number_of_reproducible_articles` articles that include R script and data files, either in the supplementary materials or in a trustworthy data repository. Figure \@ref(fig:archaeology-reproducible-articles) shows that although this number is small, it is increasing, hinting at the emergence of a new approach to how archaeologists share their research.

These archaeology articles that include publicly available R code and data are part of a new approach to scientific communication that has also been observed in other fields. Although we are not aware of attempts to quantify this change in the same way we have for archaeology, we have observed a number of recent publications that describe and recommend code sharing in statistics [@Baumer2014], genome biology [@markowetz2015five], computational biology [@Sandve2013], biostatistics [@Peng2009], computer science [@peng2011reproducible; @Mitchell2012], applied mathematics [@LeVeque2012], speech science [@Abari2012], political science [@king1995replication; Dafoe2014], and the social sciences generally [@Miguel2014]. As part of this growing interest in using code for research we also see manifestos aimed at researchers doing any kind of quantitative work [@nosek2015promoting; @Ince2012; @Barnes2010] and articles recommending using R in undergraduate education [@Bray2014; @Eglen2009]. Much of this literature is concerned with identifying and solving problems of irreproducibility in research. This comes from a renewed interest in basic principles of the scientific method: that once-off results should not be trusted, instead they should be reproducible by other members of the research community. Furthermore, an article should sufficiently describe the presented research results such that a colleague can fully understand the results and how they were obtained. Many of these authors trace problems of irreproducibility to the increased complextity of computer-based analyses (but only a limited space to describe them in a journal article), and the use of mouse-driven computer programs where the researcher's analytical decisions are not recorded during the research process, and thus cannot be shared with other researchers. In our view, using R (or similar open source programming languages) to conduct research might not constitute a tool-driven revolution, but publicaly sharing the code used for research, as a solution to problems of irreproducibility could enable revolutionary change, since openly shared code could speed the adoption of code throughout a research community.
